A Walk In The Classroom - Using Fitbit Charge 4 to measure expert and novice teachers' heart rate and step counts in a laboratory study

### 2022-09-12 02_paper/02_study/02_fitbit paper/fitbit_paper.Rmd
fitbit paper

needs(tidyverse,
      lubridate,
      viridis,
      grid,
      gridExtra,
      cowplot,
      readxl,
      ARTofR,
      moments,
      ltm,
      sjPlot,
      kableExtra,
      xtable,
      DescTools,
      formattable)

@@

>>>>

needs(tidyverse,
      lubridate,
      janitor,
      viridis,
      grid,
      gridExtra,
      cowplot,
      readxl,
      ARTofR,
      moments,
      ltm,
      sjPlot,
      kableExtra,
      xtable,
      DescTools,
      formattable)

```
# Introduction
Especially the first five years are a crucial time the teaching profession, whether the development from novice to expert is experienced as successful and as an increase in competence or leads to the fact that between 40 and 50 percent of the beginning teachers left the profession [@ingersoll2003wrong; @prilleltensky2016teacher]. One reason among many why teachers end their careers is the demanding workload and resulting stress in the teaching profession [@struyven2014teachers]. In her study with 400 secondary teachers, @fisher2011factors showed that burnout scores were significantly different between novice and experienced teachers, whereby novices had higher burnout. Furthermore, she found that low years of experience was a statistically significant predictor for stress in the teaching profession, along with job satisfaction and burnout.

XXX ÜBERGANG VON STRESS ZU HEART RATE XXX

The heart rate is considered an indicator of physical or emotional stress. Increased workload is accompanied by increased heart rate. Therefore, heart rate can be considered the most important psycho physiological parameter [@sachs2014physiologische].

XXX ÜBERGANG VON HEART RATE ZU SMARTWATCHES - VORTEILE XXX low-cost, non-invasive, wrist-worn devices

There is a gap in research on whether heart rate measurements using fitness watches are appropriate to detect differences between experienced and inexperienced teachers in a laboratory setting.

## Aim of the study
The aim of the study was to uncover differences between 25 experienced and 25 inexperienced teachers on the physical and emotional demands during a 15-minute micro-teaching unit by measuring their heart rate using a fitness watch (Fitbit Charge 4). As the fitness watch was put on at the beginning of the survey and was not taken off until the entire data collection was completed, the heart rate could be measured in the resting and arousal states. In addition, the watch recorded the subjects' movement activity to determine how many steps were taken during the lesson. The practicality of the FitBit for classroom research is discussed.

@@

### FitBit

All participants were given a FitBit Smart Watch Charge 4 to wear during the experiment.

## Heart Rate

The heart rate of each subject were measured before, during and after the experiment.

XXX Data and data wrangling need to be imported XXX 
```{r}
# # Initial Preparations
# ## Run following Code if and only if new data is added to the initial data.
# 
# source(file = "data.wrangling.R")
# 
# 
# # Plots
# 
# df_heart_all <-
#   read_rds(file = "final_data/heart_all.rds") |> 
#   mutate(status = factor(status,
#                          levels = c("Expert","Novice"),
#                          labels = c("Expert","Novice")
#                          )
#          )
# 
# df_heart_min <-
#   read_rds(file = "final_data/heart_min.rds")
# 
# df_heart_aggr <- 
#   read_rds(file = "final_data/heart_aggr.rds") |> 
#   mutate(status = factor(status,
#                          levels = c("Novice","Expert"),
#                          labels = c("Novice","Expert")
#                          )
#          )
# 
# plot_flow <-
# ggplot(data = df_heart_all,
#        mapping = aes(x = as.numeric(Time),
#                      y = `Heart Rate`,
#                      group = ID)
#        ) +
#   geom_line(mapping = aes(color = status),
#             size = 0.3,
#             alpha = 1
#             ) +
#   xlim(0, min(df_heart_min$min)) +
#   scale_color_viridis_d(option = "A",
#                        end = 0.8)  +
#   scale_linetype_manual(values = c(1, 6)) +
#   labs(x = "Time (in Seconds)",
#        y = "Heart Rate\n(in Beats per Minute)") +
#   theme_minimal() +
#   theme(text = element_text(family = "serif"),
#         legend.position = "none",
#         strip.text = element_text(size = 16)) +
#   facet_grid(cols = vars(Type),
#              scales = "free_x")
# 
# plot_flow_leg <-
# ggplot(data = df_heart_all,
#        mapping = aes(x = Time,
#                      y = `Heart Rate`
#                      )
#        ) +
#   geom_smooth(mapping = aes(color = status),
#               method = "loess") +
#   scale_color_viridis_d(option = "A",
#                         end = 0.8) +
#  # scale_linetype_manual(values = c(1, 6)) +
#   theme_minimal() +
#   theme(text = element_text(family = "serif"),
#         legend.title = element_blank()) 
# 
# legend <- cowplot::get_legend(plot_flow_leg)
# 
# plot_loess <-
# ggplot(data = filter(df_heart_all),
#        mapping = aes(x = as.numeric(Time),
#                      y = `Heart Rate`
#                      )
#        ) +
#   stat_smooth(geom = "line",
#               mapping = aes(color = status),
#               method = "loess",
#               se = FALSE) +
#   xlim(0,min(df_heart_min$min)) +
#   scale_color_viridis_d(option = "A",
#                         end = 0.8) +
#   scale_linetype_manual(values = c(1, 6)) +
#   labs(x = "Time (in Seconds)",
#        y = "Heart Rate\n(in Beats per Minute)") +
#   theme_minimal() +
#   facet_grid(cols = vars(Type),
#              scales = "free_x",
#              ) +
#   theme(text = element_text(family = "serif"),
#         legend.position = "none",
#         strip.background = element_blank(),
#         strip.text.x = element_blank()
#         )
# 
# plot_box <- 
# ggplot(data = df_heart_aggr,
#        mapping = aes(x = Type,
#                      y = mean,
#                      color = status
#                      )
#        ) +
#   geom_boxplot(color = "darkgrey",
#                fill = "grey") +
#   geom_line(data = filter(df_heart_aggr,
#                           status == "Novice"),
#             mapping = aes(x = Type,
#                           y = mean,
#                           group = as_factor(ID)
#                           ),
#             linetype = 1,
#             alpha = 0.5
#             ) +
#   geom_line(data = filter(df_heart_aggr,
#                           status == "Expert"),
#             mapping = aes(x = Type,
#                           y = mean,
#                           group = as_factor(ID)
#                           ),
#             linetype = 6,
#             alpha = 0.5
#             ) +
#   geom_point(size = 2,
#              alpha = 0.5) +
#   labs(x = "",
#        y = "Average Heart Rate\n(in Beats per Minute)") +
#   scale_color_viridis_d(option = "A",
#                        end = 0.8) +
#   theme_minimal() +
#   theme(text = element_text(family = "serif"),
#         strip.text = element_blank(),
#        legend.position = "none",
#         axis.text.x = element_blank()
#         ) +
#   aes(fill = Type)
# 
# lo_mat <- rbind(c(1,1,1,1,1,1,NA),
#                 c(2,2,2,2,2,2,4),
#                 c(3,3,3,3,3,3,NA))
# 
# plot_heart <- 
# grid.arrange(grobs = list(plot_flow, plot_loess, plot_box, legend),
#              layout_matrix = lo_mat)
# 
# plot_heart

### Steps

Using the FitBit we measured the steps a person walked during the experiment as well. We validated the steps in six randomly selected videos by manual step counting. A step was considered to be the movement of the subject's foot forward, backward, or sideways. The measurement was not as reliable as hoped.

```{r}
# df_steps <- read_rds(file = "final_data/steps.rds")
# set.seed(5678)
# ggplot(data = df_steps,
#        mapping = aes(x = type,
#                      y = Steps
#                      )
#        ) +
#   geom_boxplot(mapping = aes(color = type),
#                fill = "grey",
#                alpha = 0.6) +
#   geom_jitter(mapping = aes(color = type),
#               alpha = 0.9,
#               width = 0.3) + 
#   scale_color_viridis_d(option = "A",
#                        end = 0.8) +
#   labs(x = "",
#        y = "Steps During\nMain Phase") +
#   theme_minimal() +
#   theme(text = element_text(family = "serif"),
#         legend.title = element_blank(),
#         axis.text.x = element_blank(),
#         axis.ticks.x = element_blank()
#         ) 
# 
# ### Validation of steps 
# # selecting some rows at random with sample_n()
# sample_n(df_steps, 6)
# 
# # random ID Output:
# # Novices ID: 102, 125, 110
# # Experts ID: 219, 218, 214
# 
# # Results from manual step counting:
# # Novices ID: 110 -> 178 steps; 102 -> 219 steps; 125 -> 319 steps
# # Experts ID: 214 -> 437 steps; 218 -> 195 steps; 219 -> 530 steps

### 2022-09-12 02_paper/02_study/02_fitbit paper/fitbit_paper.tex
fitbit paper

@@

We used a Fitbit Charge 4 to measure the teachers' heart rate. The device was attached 2-finger widths above the ulnar styloid process to the partipant's wrist following the manufacturer's instructions. To determine the heart rate, the Fitbit flashes green LEDs many times per second and uses light-sensitive photodiodes to measure the volume changes in the capillaries and then calculates how many times the heart beats per minute (bpm). Data were automatically wireless synced with an iPad via Bluetooth to a Fitbit account, and subsequently, the intraday second-by-second data were exported for each session using the opensource software Pulse Watch (PulseWatch. URL: \url{https://iccir919.github.io/pulseWatch/public/index.html} {[}accessed 2022-08-03{]}).

@@

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

@@

@@

@@

>>>>

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Especially the first five years are a crucial time the teaching profession, whether the development from novice to expert is experienced as successful and as an increase in competence or leads to the fact that between 40 and 50 percent of the beginning teachers left the profession (Ingersoll \& Smith, 2003; Prilleltensky, Neff, \& Bessell, 2016). One reason among many why teachers end their careers is the demanding workload and resulting stress in the teaching profession (Struyven \& Vanthournout, 2014). In her study with 400 secondary teachers, Fisher (2011) showed that burnout scores were significantly different between novice and experienced teachers, whereby novices had higher burnout. Furthermore, she found that low years of experience was a statistically significant predictor for stress in the teaching profession, along with job satisfaction and burnout.

XXX ÜBERGANG VON STRESS ZU HEART RATE XXX

The heart rate is considered an indicator of physical or emotional stress. Increased workload is accompanied by increased heart rate. Therefore, heart rate can be considered the most important psycho physiological parameter (Sachs, 2014).

XXX ÜBERGANG VON HEART RATE ZU SMARTWATCHES - VORTEILE XXX low-cost, non-invasive, wrist-worn devices

There is a gap in research on whether heart rate measurements using fitness watches are appropriate to detect differences between experienced and inexperienced teachers in a laboratory setting.

\hypertarget{aim-of-the-study}{%
\subsection{Aim of the study}\label{aim-of-the-study}}

The aim of the study was to uncover differences between 25 experienced and 25 inexperienced teachers on the physical and emotional demands during a 15-minute micro-teaching unit by measuring their heart rate using a fitness watch (Fitbit Charge 4). As the fitness watch was put on at the beginning of the survey and was not taken off until the entire data collection was completed, the heart rate could be measured in the resting and arousal states. In addition, the watch recorded the subjects' movement activity to determine how many steps were taken during the lesson. The practicality of the FitBit for classroom research is discussed.

@@

We used a Fitbit Charge 4 to measure the teachers' heart rate. The device was attached 2-finger widths above the ulnar styloid process to the subject's wrist following the manufacturer's instructions. To determine the heart rate, the Fitbit flashes green LEDs many times per second and uses light-sensitive photodiodes to measure the volume changes in the capillaries and then calculates how many times the heart beats per minute (bpm). Data were automatically wireless synced with an iPad via Bluetooth to a Fitbit account, and subsequently, the intraday second-by-second data were exported for each session using the opensource software Pulse Watch (PulseWatch. URL: \url{https://iccir919.github.io/pulseWatch/public/index.html} {[}accessed 2022-08-03{]}).

@@

\hypertarget{fitbit}{%
\subsubsection{FitBit}\label{fitbit}}

All participants were given a FitBit Smart Watch Charge 4 to wear during the experiment.

\hypertarget{heart-rate}{%
\subsection{Heart Rate}\label{heart-rate}}

The heart rate of each subject were measured before, during and after the experiment.

XXX Data and data wrangling need to be imported XXX

\hypertarget{steps}{%
\subsubsection{Steps}\label{steps}}

Using the FitBit we measured the steps a person walked during the experiment as well. We validated the steps in six randomly selected videos by manual step counting. A step was considered to be the movement of the subject's foot forward, backward, or sideways. The measurement was not as reliable as hoped.

@@

\hypertarget{discussion-and-implication}{%
\section{Discussion and implication}\label{discussion-and-implication}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-fisher2011factors}{}}%
Fisher, M. H. (2011). Factors influencing stress, burnout, and retention of secondary teachers. \emph{Current Issues in Education}, \emph{14}(1).

@@

\leavevmode\vadjust pre{\hypertarget{ref-ingersoll2003wrong}{}}%
Ingersoll, R. M., \& Smith, T. M. (2003). The wrong solution to the teacher shortage. \emph{Educational Leadership}, \emph{60}(8), 30--33.

@@

\leavevmode\vadjust pre{\hypertarget{ref-prilleltensky2016teacher}{}}%
Prilleltensky, I., Neff, M., \& Bessell, A. (2016). Teacher stress: What it is, why it's important, how it can be alleviated. \emph{Theory Into Practice}, \emph{55}(2), 104--111.

@@

\leavevmode\vadjust pre{\hypertarget{ref-sachs2014physiologische}{}}%
Sachs, S. (2014). \emph{Physiologische parameter zur bewertung der lernwirksamkeit von lernsituationen} (PhD thesis). Universit{ä}t Ulm.

\leavevmode\vadjust pre{\hypertarget{ref-struyven2014teachers}{}}%
Struyven, K., \& Vanthournout, G. (2014). Teachers' exit decisions: An investigation into the reasons why newly qualified teachers fail to enter the teaching profession or why those who do enter do not continue teaching. \emph{Teaching and Teacher Education}, \emph{43}, 37--45.
